---
title: PSet example
author: Jozef Rivest
date: last-modified

format:
  pset-598-599-pdf: default

bibliography: references.bib
---

## Problem 1

### (a)

To find the Hessian matrix from the following equation: $f(x,y) =
3x^4 + 4y^2 - 3xy + 12$, we first have to find the derivative vector
containing the first partial derivatives.

```{=tex}
\begin{equation}
\mathbf{D}(f(x,y)) = 
\begin{bmatrix}
12x^2 - 3y & 8y - 3x
\end{bmatrix}
\end{equation}
```

Then, we compute the second own an cross-partial derivatives and fill 
the Hessian. The Equation \ref{eq:Hessian} shows the result.

```{=tex}
\begin{equation}\label{eq:Hessian}
\mathbf{H}(f(x,y)) = 
\begin{bmatrix}
24x & -3 \\
-3 & 8
\end{bmatrix}
\end{equation}
```

This is similar to what @moore_siegel13 explain in their book. 

## Problem 2 

### (a)

To generate the sample using the inverse-CDF method, we can rely on 
the *universality of the uniform distribution* to answe this question. 
Using the fact that the distribution of the percentile is uniformally 
distributed, and by plugin these values into the quantile function, 
for this problem, we can then approximate the PDF of the normal 
distribution [@blitzstein_hwang19, p. 225-226]. 

```{r}
#| fig-cap: Histogram of generated numbers
#| fig-height: 4
# setting seed for reproductibility
set.seed(48104)

# specifying the nb of observations
n <- 10000

# generating the normal distribution
n_1 <- rnorm(n, mean = 0, sd = 1)

# plotting the result
hist(n_1)
```

```{r}
#| fig-cap: Histogram of the sample
#| fig-height: 4

# generating random numbers from uniform distribution
nb <- runif(n)

# sample using the inverse CDF
sample <- qnorm(nb, mean = 0, sd = 1)

# ploting the output
hist(sample)
```

### (b)

```{r}
#| fig-cap: Q-Q Plot
#| fig-height: 4
#| label: fig-qqplot
#| fig-pos: H

# Making the Q-Q plot
qqplot(sample, n_1)
abline(0, 1, col = "red")
```

The @fig-qqplot shows the Q-Q plot using the distribution of the
normally distributed numbers generated from `rnorm`, and the sample
generated using the inverse CDF from `qnorm`.  Q-Q plot allows to
compare two probability distribution by plotting their quantiles^[To
learn more about the Q-Q plot, you can refer to the definition and
the information provided by the Wikiepdia page:
\url{https://en.wikipedia.org/wiki/Q–Q_plot}].  As all the points
fall roughly on the same line, we can say that the point generated
with `rnorm` come from the same distribution as those generated by
`qnorm`. This makes sense given that we used the percentiles generated
from the uniform distribution and passed them to the quantile
function of the normal distribution.

{{< pagebreak >}}

# References

::: {#refs}
:::
